---
title: "Nonparametric Statistics"
author: "Noé Debrois and ID number"
date: "2024-01-12"
output:
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r,echo=FALSE,results='hide',warning=FALSE,message=FALSE}
#Library imports 
library(aplpack)
library(boot)
library(broom)
library(bslib)
library(colorspace)
library(conformalInference)
library(conformalInference.fd)
library(conformalInference.multi)
library(corrplot)
library(cowplot)
library(datasets)
library(dbscan)
library(DepthProc)
library(desc)
library(DescTools)
library(exactRankTests)
library(fastmap)
library(fda)
library(fds)
library(gam)
library(gdtools)
library(ggnewscale)
library(ggplot2)
library(ggsignif)
library(ggtext)
library(graphics)
library(grDevices)
library(grid)
library(gridExtra)
library(lme4)
library(lmPerm)
library(locfit)
library(Matrix)
library(MatrixModels)
library(mgcv)
library(nlme)
library(np)
library(npreg)
library(pbapply)
library(permute)
library(polynom)
library(pyinit)
library(quantreg)
library(rainbow)
library(Rcpp)
library(resample)
library(roahd)
library(RobStatTM)
library(robustbase)
library(rpart)
library(rrcov)
library(scales)
library(sm)
library(splines)
library(stats)
library(survival)
library(survminer)

library(base64enc)
library(bit)
library(brio)
library(cachem)
library(commonmark)
library(cpp11)
library(doParallel)
library(parallel)
library(parallelly)

```


# Exercise 1

# A friend of Andrew O’Cappor, Dr. Alexander House, statistician and chemometrician, advises the farmer that spectroscopy is the state-of-the-art technology to employ when it comes to evaluate milk quality. Motivated by this, and advised by Dr House, Andrew O’Cappor is interested in building a nonparametric model to predict Milk pH by means of the absorbance values at wavenumbers 70 and 300 cm-1, contained in the df_2.Rds file. He therefore asks you to:

## Question 1.1

# 1. Build a degree 2 b-spline model to regress Milk pH on the milk absorbance at wavenumber 70cm-1. Set knots at the first and third quartiles of the explanatory variable. Provide a plot of the regression line with standard errors for the prediction, a table summarizing the coefficients and comment the results.

**Synthetic description of assumptions, methods, and algorithms **    


**Results and brief discussion**    

```{r,echo=FALSE}
df_2 <- readRDS("/Users/noedebrois/Desktop/Politecnico/Non Parametric Statistics/LECTURES/PREVIOUS EXAMS-20231223/Data/2023-07-06/df_2.Rds")
```

```{r}
knots <- quantile(df_2$wave_70,probs = c(.25, .75))

fit_spline <- lm(Native_pH~bs(wave_70,knots = knots,degree = 2), data = df_2)

knitr::kable(broom::tidy(summary(fit_spline)))
```
```{r}
new_data_seq <- seq(min(df_2$wave_70), max(df_2$wave_70), length.out = 100)

preds=predict(fit_spline, newdata = list(wave_70=new_data_seq),se=T)
se.bands=cbind(preds$fit +2* preds$se.fit ,preds$fit -2* preds$se.fit)

plot(y=df_2$Native_pH,x=df_2$wave_70 ,cex =.5, col =" darkgrey ")
lines(new_data_seq,preds$fit ,lwd =2, col =" blue")
matlines(new_data_seq, se.bands ,lwd =1, col =" blue",lty =3)

knots_pred=predict(fit_spline,list(wave_70=knots))
points(knots,knots_pred, col='blue',pch=19)
abline(v = knots, lty=3, col="blue")
```

## Question 1.2

# 2. Build an additive model for regressing Milk pH on the milk absorbance at wavenumbers 70cm-1 and 300cm-1 using penalized regression splines. Write in proper mathematical form the additive model you estimate, report the adjusted R2 and the parametric p-values of the tests. Comment on the results.

**Synthetic description of assumptions, methods, and algorithms **    
abc     

**Results and brief discussion**    
123
```{r,echo=FALSE}
fit_gam <- gam(Native_pH ~ s(wave_70, bs = "cr") + s(wave_300, bs = "cr"), data = df_2)
table_fit_gam <- summary(fit_gam)
(r_2_squared <- table_fit_gam$r.sq)
```
```{r}
table_fit_gam$p.pv
```
```{r}
table_fit_gam$s.table
```


## Question 1.3

# 3. Build a semi-parametric model for regressing Milk pH on the milk absorbance at wavenumbers 70cm-1 and 300cm-1, assuming the effect of wavenumber 300cm-1 to be linear on the response. Using a bootstrap approach, provide a reverse percentile confidence interval for the parametric effect of level 95%

**Synthetic description of assumptions, methods, and algorithms **    
abc     

**Results and brief discussion**    
123
```{r,echo=FALSE}
fit_semi <- gam(Native_pH~s(wave_70,bs = "cr") + wave_300,data = df_2)

fitted.obs <- fit_semi$fitted.values
res.obs <- fit_semi$residuals

wave300_obs = summary(fit_semi)$p.table[2,1]
B = 1000
T.boot.wave300 = numeric(B)
set.seed(2022)
for(b in 1:B) {
  Native_pH_boot <- fitted.obs + sample(res.obs, replace = T) 
  fit_semi_boot <- gam(Native_pH_boot ~ s(df_2$wave_70, bs = "cr") + df_2$wave_300)
  fit_semi_boot_table = summary(fit_semi_boot)
  T.boot.wave300[b] = fit_semi_boot_table$p.table[2,1]
}

alpha <- 0.05

right.quantile.wave300 <- quantile(T.boot.wave300, 1 - alpha/2)
left.quantile.wave300 <- quantile(T.boot.wave300, alpha/2)

CI.RP.wave300 <- c(wave300_obs - (right.quantile.wave300 - wave300_obs),
                   wave300_obs - (left.quantile.wave300- wave300_obs)) 

names(CI.RP.wave300)=c('lwr','upr')

CI.RP.wave300

```

