---
title: "Nonparametric Statistics"
author: "Noé Debrois, Person Code 10949145, ID 242751"
date: "2024-01-12"
output:
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r,echo=FALSE,results='hide',warning=FALSE,message=FALSE}
#Library imports 
library(aplpack)
library(boot)
library(broom)
library(bslib)
library(colorspace)
library(conformalInference)
library(conformalInference.fd)
library(conformalInference.multi)
library(corrplot)
library(cowplot)
library(datasets)
library(dbscan)
library(DepthProc)
library(desc)
library(DescTools)
library(exactRankTests)
library(fastmap)
library(fda)
library(fds)
library(gam)
library(gdtools)
library(ggnewscale)
library(ggplot2)
library(ggsignif)
library(ggtext)
library(graphics)
library(grDevices)
library(grid)
library(gridExtra)
library(lme4)
library(lmPerm)
library(locfit)
library(Matrix)
library(MatrixModels)
library(mgcv)
library(nlme)
library(np)
library(npreg)
library(pbapply)
library(permute)
library(polynom)
library(pyinit)
library(quantreg)
library(rainbow)
library(Rcpp)
library(resample)
library(roahd)
library(RobStatTM)
library(robustbase)
library(rpart)
library(rrcov)
library(scales)
library(sm)
library(splines)
library(stats)
library(survival)
library(survminer)

library(base64enc)
library(bit)
library(brio)
library(cachem)
library(commonmark)
library(cpp11)
library(doParallel)
library(parallel)
library(parallelly)

```


# Exercise 3

## Question 3.1
**Synthetic description of assumptions, methods, and algorithms **    
We will use functional data (using roahd package). Then if we defined well our functional data, we can just use plot(f_data). After having plotted our functional boxplot, we can have access to the magnitude outliers using $ID_outliers argument.

**Results and brief discussion**    

```{r,echo=FALSE}
ex3 <- readRDS("/Users/noedebrois/Desktop/Politecnico/Non Parametric Statistics/Exam/NPS Exam 12_01_2024/NPS_January24/ex3.rds")
```

```{r}
grid <- ex3$time
df3 = ex3$df3
df3 = df3[,3:102]
f_data <- fData(grid, df3)
```

```{r, fig.width=4}
plot(f_data)
```
```{r, fig.width=4}
fbplot <- fbplot(f_data, main="Magnitude outliers")
```
```{r}
outliers = c(fbplot$ID_outliers) 
outliers # yes there is one !
```
```{r}
ex3$df3[outliers,] # The magnitude outlier is Luxembourg !
```

## Question 3.2
**Synthetic description of assumptions, methods, and algorithms **    
We use here the trim mean which is a robustified version of the classical mean.

**Results and brief discussion**    
123
```{r,echo=FALSE}
df_clean_group = split(x = ex3$df3, f = ex3$df3$IncomeGroup)
first_group = df_clean_group$`High income`[,-1]
first_group = first_group[,-1]
B = nrow(first_group)
mean_trim <- numeric(B)
for (b in 1:B) {
  a = mean(first_group[,b], trim = .1)
  mean_trim = c(mean_trim, a)
}
mean_trim
```


## Question 3.3
**Synthetic description of assumptions, methods, and algorithms **    
abc     

**Results and brief discussion**    
123
```{r,echo=FALSE}

```

## Question 3.4
**Synthetic description of assumptions, methods, and algorithms **    
Assumptions : X11,...,X1n~p1 X21,...,X2m~p2
H0 : p1=p2 , H1 : p1!=p2 
T0 = |X_trim_mean1-X_trim_mean2|

**Results and brief discussion**    
123
```{r,echo=FALSE}

```

```{r,echo=FALSE}

```


