---
title: "Nonparametric Statistics"
author: "Noé Debrois, Person Code 10949145, ID 242751"
date: "2024-01-12"
output:
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r,echo=FALSE,results='hide',warning=FALSE,message=FALSE}
#Library imports 
library(aplpack)
library(boot)
library(broom)
library(bslib)
library(colorspace)
library(conformalInference)
library(conformalInference.fd)
library(conformalInference.multi)
library(corrplot)
library(cowplot)
library(datasets)
library(dbscan)
library(DepthProc)
library(desc)
library(DescTools)
library(exactRankTests)
library(fastmap)
library(fda)
library(fds)
library(gam)
library(gdtools)
library(ggnewscale)
library(ggplot2)
library(ggsignif)
library(ggtext)
library(graphics)
library(grDevices)
library(grid)
library(gridExtra)
library(lme4)
library(lmPerm)
library(locfit)
library(Matrix)
library(MatrixModels)
library(mgcv)
library(nlme)
library(np)
library(npreg)
library(pbapply)
library(permute)
library(polynom)
library(pyinit)
library(quantreg)
library(rainbow)
library(Rcpp)
library(resample)
library(roahd)
library(RobStatTM)
library(robustbase)
library(rpart)
library(rrcov)
library(scales)
library(sm)
library(splines)
library(stats)
library(survival)
library(survminer)

library(base64enc)
library(bit)
library(brio)
library(cachem)
library(commonmark)
library(cpp11)
library(doParallel)
library(parallel)
library(parallelly)

```


# Exercise 2

## Question 2.1
**Synthetic description of assumptions, methods, and algorithms **    
We apply a raw polynomial fit here twice.
On the first, we see thanks to the pvalues that all the coefficient are *** significant. 

**Results and brief discussion**    

```{r,echo=FALSE}
df2 <- read.csv("/Users/noedebrois/Desktop/Politecnico/Non Parametric Statistics/Exam/NPS_January24/df2.csv", header=TRUE, sep=",")
```

```{r}
m_quad_1=lm(df2$CO2_emissions ~ I(df2$GDP_per_capita^2))
summary(m_quad_1)
```
```{r}
gdp.grid=seq(range(df2$GDP_per_capita)[1],range(df2$GDP_per_capita)[2],by=1410)
preds=predict(m_quad_1,list(gdp=gdp.grid),se=T)
se.bands=cbind(preds$fit +2* preds$se.fit ,preds$fit -2* preds$se.fit)

plot(df2$GDP_per_capita, df2$CO2_emissions, xlim=range(gdp.grid) ,cex =.5, col =" darkgrey ", main='Quadratic Fit')
lines(gdp.grid, preds$fit,lwd =2, col =" blue") 
matlines (gdp.grid ,se.bands ,lwd =1, col =" blue",lty =3)
```
```{r}
gdp.grid=seq(range(df2$GDP_per_capita)[1],range(df2$GDP_per_capita)[2],by=1410)
preds=predict(m_quad_1,list(gdp=gdp.grid),se=T)
se.bands=cbind(preds$fit +2* preds$se.fit ,preds$fit -2* preds$se.fit)

plot(df2$GDP_per_capita, df2$CO2_emissions, xlim=range(gdp.grid) ,cex =.5, col =" darkgrey ", main='Quadratic Fit')
lines(gdp.grid, preds$fit,lwd =2, col =" blue") 
matlines (gdp.grid ,se.bands ,lwd =1, col =" blue",lty =3)
```

## Question 2.2
**Synthetic description of assumptions, methods, and algorithms **    
Assumptions : X1,...,Xn~p
We want to predict the location of Xn+1 following the same distribution p. In particular, we want to build a confidence region for this new point.
We choose a Non-conformity measure. Then, from this non conformity measure, we can now rank the observations of an augmented vector (x1,...,xn,xtest) and so determine the p-value corresponding to xtest : pval=R_{n+1}/(n+1).
From that we can deduce prediction bands or prediction region with confidence 1-alpha by taking C(X1,...,Xn)={xn+1 in R^p st r_{n+1}<=ceil((1-alpha)*(n+1))}

**Results and brief discussion**    
123
```{r,echo=FALSE}
df2_1 = df2[, 4:5]
m_quad1 = lm(df2_1$CO2_emissions ~ I(df2_1$GDP_per_capita))
summary(m_quad1)
```


## Question 2.3
**Synthetic description of assumptions, methods, and algorithms **    
By fitting the GAM and looking at the p-values, we see that only the intercept and the second coefficient from the bs are statistically significant (others have higher than 0.05 pvalues).
The coefficients represent the estimated effects of the basis splines for the 
variable GDP. The asterisks in the p-values indicate statistical significance.
The pvalues < 0.05 suggest that the related coefficients contribute significantly to the model.
The others suggest that these terms may not be contributing significantly to the model.

**Results and brief discussion**    
123
```{r,echo=FALSE}
df = df2
df[,6] = log(df[,6]/(1-df[,6]))

fit_gam <- gam(Gini ~ bs(GDP_per_capita, degree = 2, knots = median(3000)), 
               data = df)
summary(fit_gam) 
```

## Question 2.4
**Synthetic description of assumptions, methods, and algorithms **    
Bootstrap Assumptions :
S1 = (X1,...,Xn)~ iid p
The primary task of bootstrapping is estimating from a random sample S1, the distribution of a statistic, it is to say an estimator theta_hat of an unknown parameter theta.
The principle is to sample a new vector having the same size as S1 from S1 itself with replacement ! From this idea, we can provide more information about the quality of the estimator.  

**Results and brief discussion**    
123
```{r,echo=FALSE}

```

```{r,echo=FALSE}

```


