---
title: "Nonparametric Statistics"
author: "Noé Debrois, Person Code 10949145, ID 242751"
date: "2024-01-12"
output:
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r,echo=FALSE,results='hide',warning=FALSE,message=FALSE}
#Library imports 
library(aplpack)
library(boot)
library(broom)
library(bslib)
library(colorspace)
library(conformalInference)
library(conformalInference.fd)
library(conformalInference.multi)
library(corrplot)
library(cowplot)
library(datasets)
library(dbscan)
library(DepthProc)
library(desc)
library(DescTools)
library(exactRankTests)
library(fastmap)
library(fda)
library(fds)
library(gam)
library(gdtools)
library(ggnewscale)
library(ggplot2)
library(ggsignif)
library(ggtext)
library(graphics)
library(grDevices)
library(grid)
library(gridExtra)
library(lme4)
library(lmPerm)
library(locfit)
library(Matrix)
library(MatrixModels)
library(mgcv)
library(nlme)
library(np)
library(npreg)
library(pbapply)
library(permute)
library(polynom)
library(pyinit)
library(quantreg)
library(rainbow)
library(Rcpp)
library(resample)
library(roahd)
library(RobStatTM)
library(robustbase)
library(rpart)
library(rrcov)
library(scales)
library(sm)
library(splines)
library(stats)
library(survival)
library(survminer)

library(base64enc)
library(bit)
library(brio)
library(cachem)
library(commonmark)
library(cpp11)
library(doParallel)
library(parallel)
library(parallelly)

```


# Exercise 1

## Question 1.1
**Synthetic description of assumptions, methods, and algorithms **    
The median is the deepest point in the sense of the Tukey depth.
ROM (Romania) has the highest tukey depth.
Then PAN.
Then MEX.
**Results and brief discussion**    

```{r,echo=FALSE}
df1 <- read.csv("/Users/noedebrois/Desktop/Politecnico/Non Parametric Statistics/Exam/NPS_January24/df1.csv", header=TRUE, sep=",")

df_red = df1[,c(1,4,5,6)]
df_red[,4] = log(df_red[,4]/(1-df_red[,4]))

tukey_median <- depthMedian(df_red[,2:4], depth_params = list(method='Tukey'))
tukey_median
```
```{r}
df_red = df_red[-53,]
tukey_median2 <- depthMedian(df_red[,2:4], depth_params = list(method='Tukey'))
tukey_median2
```
```{r}
df_red = df_red[-48,]
tukey_median3 <- depthMedian(df_red[,2:4], depth_params = list(method='Tukey'))
tukey_median3
```

## Question 1.2
**Synthetic description of assumptions, methods, and algorithms **    
We can build automatically the bagplot matrix using the aplpack library. We need a bagplot MATRIX since we have three dimensions (GDP, CO2 and Gini) and since a bagplot only compares two dimensions at a time. Then we plot only the bagplot based on the two dimensions of interest and extract outliers using the $pxy.outlier argument. We can retrieve the name of the using by looking in the df_red dataframe.

**Results and brief discussion**    
123
```{r,echo=FALSE}
bagplot_matrix <- aplpack::bagplot.pairs(df_red)
```
```{r}
df_bagplot = df_red[,c(2,3)]
bagplot_1 = aplpack::bagplot(df_bagplot,show.whiskers = F,main="Bagplot")
outlying_obs_1 <- bagplot_1$pxy.outlier
ind_outlying_obs_1 <- which(apply(df_bagplot,1,function(x) all(x %in% outlying_obs_1))) 
```
```{r}
ind_outlying_obs_1
```
```{r}
df_red[35,]
```


```{r}
df_red[46,]
```
```{r}
df_red[59,]
```
```{r}
df_red[60,]
```
```{r}
df1_clean <- df1[-ind_outlying_obs_1,]
```


## Question 1.3
**Synthetic description of assumptions, methods, and algorithms **  
Test statistics will be : Wilks Lambda.
Note that wilk's lambda is significant for small values.
It is sufficient to change its sign to use it in a permutation test. 
We are in a multivariate framework,so wee permute the labels associated to each unit.

**Results and brief discussion**    
123
```{r,echo=FALSE}
df1_clean_group = split(x = df1_clean, f = df1_clean$IncomeGroup)
df1_clean_group$`High income` = df1_clean_group$`High income`[, 4:6]
df1_clean_group$`Middle income` = df1_clean_group$`Middle income`[, 4:6]
df1_clean_group$`Low income` = df1_clean_group$`Low income`[, 4:6]


# IncomeGroup <- factor(df1_clean$IncomeGroup, labels=c('High income','Middle income','Low income'))
# fit <- manova(as.matrix(df1_clean) ~ IncomeGroup)
```

```{r}
```

## Question 1.4
**Synthetic description of assumptions, methods, and algorithms **    
We perform here a TWO-SIDED MANN-WHITNEY U-TEST. 
H0: P(CO2_high_income > CO2_middle_income)  = 0.5
H1: P(CO2_high_income > CO2_middle_income) != 0.5
The pvalue is bigger than 0.05 so we cannot reject H0.
**Results and brief discussion**    
123
```{r,echo=FALSE}
PI1 <- df1_clean[df1_clean$IncomeGroup == "High income",]
PI2 <- df1_clean[df1_clean$IncomeGroup == "Middle income",]
PI1 = PI1[, 5]
PI2 = PI2[, 5]
PI = c(PI1, PI2)
n1 <- length(PI1)
n2 <- length(PI2)
n  <- length(PI)

# Visualisation de la structure de dépendance
diff <- expand.grid(PI1, PI2)[,2] - expand.grid(PI1, PI2)[,1]
image(1:n1, 1:n2, matrix(sign(diff), n1, n2))

```
```{r}
```

```{r,echo=FALSE}
```


