---
title: "Nonparametric Statistics"
author: "Simon, Trottier and 10988912"
date: "2024-01-12"
output:
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r,results='hide',message=FALSE}
#Library imports 
library(aplpack)
library(boot)
library(broom)
library(bslib)
library(colorspace)
library(conformalInference)
library(conformalInference.fd)
library(conformalInference.multi)
library(corrplot)
library(cowplot)
library(datasets)
library(dbscan)
library(DepthProc)
library(desc)
library(DescTools)
library(exactRankTests)
library(fastmap)
library(fda)
library(fds)
library(gam)
library(gdtools)
library(ggnewscale)
library(ggplot2)
library(ggsignif)
library(ggtext)
library(graphics)
library(grDevices)
library(grid)
library(gridExtra)
library(lme4)
library(lmPerm)
library(locfit)
library(Matrix)
library(MatrixModels)
library(mgcv)
library(nlme)
library(np)
library(npreg)
library(pbapply)
library(permute)
library(polynom)
library(pyinit)
library(quantreg)
library(rainbow)
library(Rcpp)
library(resample)
library(roahd)
library(RobStatTM)
library(robustbase)
library(rpart)
library(rrcov)
library(scales)
library(sm)
library(splines)
library(stats)
library(survival)
library(survminer)

library(base64enc)
library(bit)
library(brio)
library(cachem)
library(commonmark)
library(cpp11)
library(doParallel)
library(parallel)
library(parallelly)

```


```{r}
```

# Exercise 3

## Question 3.1
**Synthetic description of assumptions, methods, and algorithms **    The principle behind functinal boxplot is the following one : 

Compute data depths, rank data and consider the 50% central
region, inflate it by a factor F >1, obtain the fences given by the envelope of functions inside the inflated regions and flag the observations crossing the fence or outside as amplitude outliers.


**Results and brief discussion**    

```{r}
```


## Question 3.2
**Synthetic description of assumptions, methods, and algorithms **    The $\alpha$=0.1-trimmed mean is correspond to the mean we obtain when we remove the 10% Highest and 10% Lowest values in our data. This allows to be more robust since we can have up to 10% of totally incoherent values (+inf for instance).

**Results and brief discussion**    


```{r}

```


## Question 3.3
**Synthetic description of assumptions, methods, and algorithms **    
The test hypothesis are 
H0 : High and low incomes distributions are the same 
H1 : High and low incomes distributions are different

We have first an initial value of the test T0 corresponding to the L2 distance between the functionals $\alpha$-trimmed means.

We assume first group as m members and second group as n.
The permutation test will, at each turn, assign randomly (without replacement) among the n+m members, m forming the new first group and the n others one will form a new second group. For these two groups, we then determine the value of the test T2_i.

the p-value corresponds to the numbers of test belonging to T2 that are more extreme that T0.

**Results and brief discussion**    

```{r}

```



