---
title: "Nonparametric Statistics"
author: "Simon, Trottier and 10988912"
date: "2024-01-12"
output:
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r,results='hide',message=FALSE}
#Library imports 
library(aplpack)
library(boot)
library(broom)
library(bslib)
library(colorspace)
library(conformalInference)
library(conformalInference.fd)
library(conformalInference.multi)
library(corrplot)
library(cowplot)
library(datasets)
library(dbscan)
library(DepthProc)
library(desc)
library(DescTools)
library(exactRankTests)
library(fastmap)
library(fda)
library(fds)
library(gam)
library(gdtools)
library(ggnewscale)
library(ggplot2)
library(ggsignif)
library(ggtext)
library(graphics)
library(grDevices)
library(grid)
library(gridExtra)
library(lme4)
library(lmPerm)
library(locfit)
library(Matrix)
library(MatrixModels)
library(mgcv)
library(nlme)
library(np)
library(npreg)
library(pbapply)
library(permute)
library(polynom)
library(pyinit)
library(quantreg)
library(rainbow)
library(Rcpp)
library(resample)
library(roahd)
library(RobStatTM)
library(robustbase)
library(rpart)
library(rrcov)
library(scales)
library(sm)
library(splines)
library(stats)
library(survival)
library(survminer)

library(base64enc)
library(bit)
library(brio)
library(cachem)
library(commonmark)
library(cpp11)
library(doParallel)
library(parallel)
library(parallelly)

```

```{r}
df_2=read.csv("NPS_January24/df2.csv")
N<-nrow(df_2)
p<-ncol(df_2)
df_2$Gini<-logit(df_2$Gini)
```

# Exercise 2

## Question 2.1
**Synthetic description of assumptions, methods, and algorithms **    

The quadratic regression is one of the simplest form of regression but it can still be usefull when we have prior informations on the expected model.

**Results and brief discussion**    

```{r}
fit_poly <-
  lm(CO2_emissions ~ poly(GDP_per_capita, degree = 2, raw = T),data = df_2)
```

```{r}
newdata = data.frame(GDP_per_capita = with(df_2, seq(
  range(GDP_per_capita)[1],
  range(GDP_per_capita)[2],
  by = 1 )))
preds_poly = predict(fit_poly,
                      list(GDP_per_capita = newdata$GDP_per_capita),
                      se = T)
se.bands_poly = cbind(preds_poly$fit + 2 * preds_poly$se.fit ,
                       preds_poly$fit - 2 * preds_poly$se.fit)

with( df_2,
  plot(
    GDP_per_capita ,
    CO2_emissions ,
    xlim = range(newdata$GDP_per_capita) ,
    cex = .5,
    col = " darkgrey ",
    main = 'Kuznets') )

lines(newdata$GDP_per_capita,
      preds_poly$fit ,
lwd = 2,
      col = " blue")

matlines (
  newdata$GDP_per_capita ,
  se.bands_poly ,
  lwd = 1,
  col = " blue",
lty = 3 )
```
```{r}
fit_poly_b <-
  lm(Gini ~ poly(GDP_per_capita, degree = 2, raw = T),data = df_2)

preds_polyb = predict(fit_poly_b,
                      list(GDP_per_capita = newdata$GDP_per_capita),
                      se = T)
se.bands_polyb = cbind(preds_polyb$fit + 2 * preds_polyb$se.fit ,
                       preds_polyb$fit - 2 * preds_polyb$se.fit)

with( df_2,
  plot(
    GDP_per_capita ,
    Gini ,
    xlim = range(newdata$GDP_per_capita) ,
    cex = .5,
    col = " darkgrey ",
    main = 'Environnemental Kuznets') )

lines(newdata$GDP_per_capita,
      preds_polyb$fit ,
lwd = 2,
      col = " blue")

matlines (
  newdata$GDP_per_capita ,
  se.bands_polyb ,
  lwd = 1,
  col = " blue",
lty = 3 )

```
The classical Kuznet curve seems to fit to the model, at least it is a parabola but the environnemental Kuznet curve doesn't seem close at all to what was expected.

## Question 2.2
**Synthetic description of assumptions, methods, and algorithms **    


**Results and brief discussion**    


```{r}
  
```


## Question 2.3
**Synthetic description of assumptions, methods, and algorithms **    
The GAM allow to consider independant effects which can be parametric or non-parametrics and to use them all at the same time. 

**Results and brief discussion**    
123
```{r}
fit_gam <- gam(Gini~bs(GDP_per_capita,degree = 2, knots = 30000) + IncomeGroup ,data = df_2)


sum<-summary(fit_gam)

sum$p.coeff

sum$p.pv

```
Here, we can see that there seem to be a link between the covariates and the estimate since the "global" pvalue is 0.02 so below 5%. But not anyone of the covariates seems to have statistical significance.

## Question 2.4
**Synthetic description of assumptions, methods, and algorithms **    
Assumptions :
S1 = $(X_1,…,X_n)$~p

The primary task of bootstrapping is estimating from a random sample S1, the distribution of a statistic, it is to say an estimator $\hat{\theta}$ of an unknown parameter $\theta$.

The principle is to sample a new vector having the same size as S1 from S1 itself with replacement ! From this idea, we can provide more information about the quality of the estimator.
    

**Results and brief discussion**    

```{r}
fit_sei<-fit_gam
pred_kaponzyi<-predict(fit_gam, newdata = data.frame("GDP_per_capita" =20000,"IncomeGroup" ="High income"))
res.obs <- fit_gam$residuals
fitted.obs <- fit_gam$fitted.values
B<-1000
boot_pred_kaponzyi = numeric(B) 
set.seed(2024)
for (b in 1:B) {
  Gini_boot <- fitted.obs + sample(res.obs, replace = T)
  df_boot <-
    data.frame(Gini_boot,
               GDP_per_capita = df_2$GDP_per_capita, IncomeGroup = df_2$IncomeGroup)
  fit_kernel_boot <-
    gam(Gini_boot~bs(GDP_per_capita,degree = 2, knots = 30000) + IncomeGroup ,data = df_boot)
  boot_pred_kaponzyi[b] = predict(fit_kernel_boot, newdata = data.frame("GDP_per_capita" =20000,"IncomeGroup" ="High income"))
                                  
}
alpha <- 0.05
right.quantile.pred_kaponzyi <- quantile(boot_pred_kaponzyi, 1 - alpha / 2)
leftquantile.pred_kaponzyi <- quantile(boot_pred_kaponzyi, alpha / 2)
CI.RP.pred_kaponzyi <-
  c(
    pred_kaponzyi - (right.quantile.pred_kaponzyi - pred_kaponzyi),
    pred_kaponzyi - (leftquantile.pred_kaponzyi - pred_kaponzyi)
  )
names(CI.RP.pred_kaponzyi) = c('lwr', 'upr')
CI.RP.pred_kaponzyi
```
Here is a reverse pourcentile interval for the logit(Gini) of a country with a GDP_per_capita of 20000 and an High income.
