---
title: "Nonparametric Statistics"
author: "Simon, Trottier and 10988912"
date: "2024-01-12"
output:
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r,results='hide',message=FALSE}
#Library imports 
library(aplpack)
library(boot)
library(broom)
library(bslib)
library(colorspace)
library(conformalInference)
library(conformalInference.fd)
library(conformalInference.multi)
library(corrplot)
library(cowplot)
library(datasets)
library(dbscan)
library(DepthProc)
library(desc)
library(DescTools)
library(exactRankTests)
library(fastmap)
library(fda)
library(fds)
library(gam)
library(gdtools)
library(ggnewscale)
library(ggplot2)
library(ggsignif)
library(ggtext)
library(graphics)
library(grDevices)
library(grid)
library(gridExtra)
library(lme4)
library(lmPerm)
library(locfit)
library(Matrix)
library(MatrixModels)
library(mgcv)
library(nlme)
library(np)
library(npreg)
library(pbapply)
library(permute)
library(polynom)
library(pyinit)
library(quantreg)
library(rainbow)
library(Rcpp)
library(resample)
library(roahd)
library(RobStatTM)
library(robustbase)
library(rpart)
library(rrcov)
library(scales)
library(sm)
library(splines)
library(stats)
library(survival)
library(survminer)

library(base64enc)
library(bit)
library(brio)
library(cachem)
library(commonmark)
library(cpp11)
library(doParallel)
library(parallel)
library(parallelly)

```

```{r}
df_1=read.csv("NPS_January24/df1.csv")
N<-nrow(df_1)
p<-ncol(df_1)
df_1[,6]<-logit(df_1[,6])
df_bis <- df_1[,4:6]
```


# Exercise 1

## Question 1.1
**Synthetic description of assumptions, methods, and algorithms **    

The Tukey depth, also call the Half-spaced depth in reason of the computational method behind it allows to assigns a score of centrality for each point in a distribution. The more central the point is, higher is its Tukey depth. It is a robust way of computing depth. 

**Results and brief discussion**    

```{r}
d<-numeric(N)
for (i in 1:N){
  d[i]<-depthTukey(as.matrix(df_bis[i,]),as.matrix(df_bis))
}
```

```{r}
i1=1
m1=0
for (i in 1:N){
  if (d[i]>m1){
    m1<-d[i]
    i1<-i
  }
}
i2=1
m2=0
for (i in 1:N){
  if (d[i]>m2 & i!=i1){
    m2<-d[i]
    i2<-i
  }
}
i3=1
m3=0
for (i in 1:N){
  if (d[i]>m3 & i!=i1 & i!=i2){
    m3<-d[i]
    i3<-i
  }
}
c(df_1[i1,1],df_1[i2,1],df_1[i3,1])
```
So here are the three countries with the highest depth. The median correspond to the first one which is the deepest ("CRI").

## Question 1.2
**Synthetic description of assumptions, methods, and algorithms **    

The bagplot matrix is a graphical tool which allows to detect outliers by considering 2 variables at the time ans showing the outlier for each pair of variables.

**Results and brief discussion**    


```{r}
bagplot_matrix <-aplpack::bagplot.pairs(df_bis)
```
```{r}
bagplot_12 <- compute.bagplot(df_bis[, 1], df_bis[, 2])

ind_bag<-
which(apply(df_bis[, 1:2], 1, function(x)
    all(x %in% bagplot_12$pxy.bag)))

ind_out_12 <-
which(apply(df_bis[, 1:2], 1, function(x)
    all(x %in% bagplot_12$pxy.outlier)))

df_1[ind_bag,1]
```
Here are the countries that are inside the bag.

```{r}
df_1[ind_out_12,1]
```
Here are the countries that are outliers.

```{r}
df_1<-df_1[-ind_out_12,]
df_bis <- df_bis[-ind_out_12,]
```

## Question 1.3
**Synthetic description of assumptions, methods, and algorithms **    
The Permutational Manova test allows to compare multivariate distribution and to test whether or not we can assume that they are identically distributed based on their mean.    

**Results and brief discussion**    



## Question 1.4
**Synthetic description of assumptions, methods, and algorithms **    
The Mann-Whitney U test is the following:
H0 : CO2 emissions of Middle income and High income are the same
H1 : CO2 emissions of Middle income and High income are different

To perform this test, we need could have use the wilcox.test function and setting the paired argument as false.


**Results and brief discussion**    
123
```{r}

```

