---
title: "Nonparametric Statistics"
author: "Noé Debrois and ID number"
date: "2024-01-12"
output:
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r,echo=FALSE,results='hide',warning=FALSE,message=FALSE}
#Library imports 
library(aplpack)
library(boot)
library(broom)
library(bslib)
library(colorspace)
library(conformalInference)
library(conformalInference.fd)
library(conformalInference.multi)
library(corrplot)
library(cowplot)
library(datasets)
library(dbscan)
library(DepthProc)
library(desc)
library(DescTools)
library(exactRankTests)
library(fastmap)
library(fda)
library(fds)
library(gam)
library(gdtools)
library(ggnewscale)
library(ggplot2)
library(ggsignif)
library(ggtext)
library(graphics)
library(grDevices)
library(grid)
library(gridExtra)
library(lme4)
library(lmPerm)
library(locfit)
library(Matrix)
library(MatrixModels)
library(mgcv)
library(nlme)
library(np)
library(npreg)
library(pbapply)
library(permute)
library(polynom)
library(pyinit)
library(quantreg)
library(rainbow)
library(Rcpp)
library(resample)
library(roahd)
library(RobStatTM)
library(robustbase)
library(rpart)
library(rrcov)
library(scales)
library(sm)
library(splines)
library(stats)
library(survival)
library(survminer)

library(base64enc)
library(bit)
library(brio)
library(cachem)
library(commonmark)
library(cpp11)
library(doParallel)
library(parallel)
library(parallelly)

```


# Exercise 1

# Andrew O’Cappor has recently become passionate about functional data analysis (FDA): Dr. Alexander House is then ready to unleash the power of the MilkoScan FT6000 milk analyzer, producing MIRS transmittance spectra in the entire mid-infrared light region (wavenumbers 70 and 300 cm-1 of the previous exercise were just two points of this region). The MIRS for the N = 370 milk samples are contained in the df_3_A.Rds file. Andrew O’Cappor would like to exploit these modern statistical methods to further analyze his milk samples. To this extent, he requires you to:

## Question 1.1

# 1. By suitably defining a functional data object, plot the resulting spectra for the N = 370 milk samples. Then, compute the median spectrum using the modified band depth and superimpose it to the previous plot.

**Synthetic description of assumptions, methods, and algorithms **    


**Results and brief discussion**    

```{r,echo=FALSE}
df_3_A <- readRDS("/Users/noedebrois/Desktop/Politecnico/Non Parametric Statistics/LECTURES/PREVIOUS EXAMS-20231223/Data/2023-07-06/df_3_A.Rds")

grid <- 1:ncol(df_3_A)
f_data <- fData(grid,df_3_A)
plot(f_data)
band_depth <- MBD(Data = f_data)
median_curve <- median_fData(fData = f_data, type = "MBD") 
plot(f_data)
lines(grid,median_curve$values)
```

## Question 1.2

# 2. Produce an outliergram and use it to identify potential outlying curves present in the sample, setting the inflation factor F equal to 1.3. Report a vector of row indexes identifying the samples that are outliers according to the procedure.

**Synthetic description of assumptions, methods, and algorithms **    
abc     

**Results and brief discussion**    
123
```{r,echo=FALSE}
outgr_plot <- outliergram(f_data,Fvalue = 1.3)
```
```{r}
FDA_out <- c(outgr_plot$ID_outliers)
FDA_out
```


## Question 1.3

# 3. Andrew O’Cappor has received 12 extra milk samples from a neighboring farmer, Matthew O’Fountain. These samples originated from O’Fountain’s cows and are contained in the df_3_B.Rds file. Matthew would like to compute the Relative Modified Band Depth of his milk with respect to O’Cappor samples1, to figure out whether the milk analyzer he uses is correctly calibrated. Report the Relative Modified Band Depth of the 12 extra milk samples, what can you conclude about the calibration of Matthew’s milk analyzer?

**Synthetic description of assumptions, methods, and algorithms **    
abc     

**Results and brief discussion**    
123
```{r,echo=FALSE}
df_3_B <- readRDS("/Users/noedebrois/Desktop/Politecnico/Non Parametric Statistics/LECTURES/PREVIOUS EXAMS-20231223/Data/2023-07-06/df_3_B.Rds")
f_data_B <- fData(grid,df_3_B)
MBD_f_data_B <- MBD_relative(Data_target = f_data_B, Data_reference = f_data)
MBD_f_data_B
```
```{r}
plot(append_fData(f_data_B,f_data))
# For sure there is a calibration problem in O'Fountain's milk analyzer
```
